---
title: "Gap analysis of arthropod traits"
author: "Florian D. Schneider"
date: "9 März 2017"
output: pdf_document
---


# Available data

As **abundance data**, I am refering to 

> Arthropods_GL_0813.csv  :  direct transfer from Martin Gossner

```{r}

arthropods_core <- read.csv2("../data/arthropods_core/Arthropods_GL_0813.csv")
predators_core <- subset(arthropods_core, Feeding_guild == "c")
herbivores_core <- subset(arthropods_core, Feeding_guild %in% c("h"))
herbivores_core$SpeciesID[herbivores_core$SpeciesID == "trigonotylus caelestialium"] <- "Trigonotylus caelestialium"

```


The data sheet contains the pooled abundance data of arthropod core project samplings in 2008 -- 2013 of all EPs in month June and September. Extra samplings on VIPs have been omitted to match the sampling of the EP plots. 

As **trait data**, I am refering to 

> ArthropodSpeciesTraits.csv   from   Gossner MM, Simons NK, Achtziger R, Blick T, Dorow WHO, Dziock F, Köhler F, Rabitsch W, Weisser WW (2015) Data from: A summary of eight traits of Coleoptera, Hemiptera, Orthoptera and Araneae, occurring in grasslands in Germany. Dryad Digital Repository. http://dx.doi.org/10.5061/dryad.53ds2


```{r}
arthropod_traits <- read.csv("../data/arthropod_traits/ArthropodSpeciesTraits.csv", sep = ";")
```


# completeness of data? 

## 1. which abundant species are missing on the trait list

I am looking for species relative abundance > 5% within their functional group on a plot (split data for herbivores and predators). 

```{r}
library(plyr)
# herbivores
summed_abundances <- plyr::ddply(herbivores_core, .(EP,Year), summarise, sum = sum(Abundance))

herbivores_core$rel_abundance <- herbivores_core$Abundance/summed_abundances$sum[match( paste0(herbivores_core$EP, herbivores_core$Year), paste0(summed_abundances$EP, summed_abundances$Year))]
  
rel_abundances <- plyr::ddply(herbivores_core, .(SpeciesID), summarise, max = max(rel_abundance), mean = mean(rel_abundance), min = min(rel_abundance), findings = length(rel_abundance)  )
  
hist(rel_abundances$max, breaks = 32)

dominant_herbivores <- subset(rel_abundances, max > 0.05)$SpeciesID

traits = c("Body_Size", "Dispersal_ability", "Stratum_use_short", "Feeding_mode", "Feeding_tissue", "Feeding_specialization", "Feeding_plant_part")

missing_herbivores <-  dominant_herbivores[!dominant_herbivores %in% arthropod_traits$SpeciesID]
  missing_herbivores <- cbind(SpeciesID = missing_herbivores, rel_abundances[match( missing_herbivores,rel_abundances$SpeciesID), c("max", "mean", "findings")])
(missing_herbivores <- missing_herbivores[rev(order(missing_herbivores$mean)),])

```

```{r}

# predators
summed_abundances <- plyr::ddply(predators_core, .(EP,Year), summarise, sum = sum(Abundance) )

predators_core$rel_abundance <- predators_core$Abundance/summed_abundances$sum[match( paste0(predators_core$EP, predators_core$Year), paste0(summed_abundances$EP, summed_abundances$Year))]
  
rel_abundances <- plyr::ddply(predators_core, .(SpeciesID), summarise, max = max(rel_abundance), mean = mean(rel_abundance), min = min(rel_abundance), findings = length(rel_abundance) )
  
hist(rel_abundances$max, breaks = 32)

dominant_predators <- subset(rel_abundances, max > 0.05)$SpeciesID

traits = c("Body_Size", "Dispersal_ability", "Stratum_use_short", "Feeding_mode")


missing_predators <-  dominant_predators[!dominant_predators %in% arthropod_traits$SpeciesID]
  missing_predators <- cbind(SpeciesID = missing_predators, rel_abundances[match( missing_predators,rel_abundances$SpeciesID), c("max", "mean", "findings")])
(missing_predators <- missing_predators[rev(order(missing_predators$mean)),])



```




